---
title: "Supplementary material for: Brief increases in corticosterone result in immediate and lasting changes to DNA methylation in a wild bird"
author: 
  - Conor C. Taff
  - Sabrina M. McNew
  - Leonardo Campagna
  - Maren N. Vitousek
header-includes:
  - \usepackage[font={footnotesize}, labelfont={bf}]{caption}
fontsize: 12pt
output: 
  bookdown::pdf_document2:
      number_sections: FALSE
      toc: FALSE
      extra_dependencies: ["flafter"]
  bookdown::word_document2:
      number_sections: FALSE
      toc: FALSE
      extra_dependcies: ["flafter"]
  bookdown::html_document2:
      number_sections: FALSE
      toc: TRUE
bibliography: references.bib
csl: molecular-ecology.csl
---

\raggedright

```{r setup, include = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# these go under header-includes in yaml if wanted
#  - \usepackage{setspace}\doublespacing
#  - \usepackage{lineno}\linenumbers
```



```{r seq-summary}

```


Insert alignment figure. adfadsfasdfasdf adfads adfas dfasdf asdf asdf asdf asdf as fasd fa sdf asdf asd fasd f asdf asd fas df asdf asdf asdf asd fas df asdf asdf asd fas df asdf asdf asd f asdf asdf asd fasd fas df asdf asdf as dfas fa sfd asfdas adsfffffffffffffffffffffffffffffffffffffff adsf adsfadsf 

Changes in DNA methylation in response to prior challenges represent one mechanism that could play a role in the calibration of stress response systems. Epigenetic modification by DNA methylation can alter phenotypes by making genes or promoters more or less accessible for transcription [cite]. It is well known that early life experiences can have profound programming effects on DNA methylation patterns that often persist throughout the lifetime [cite]. For example, classic work in lab rodents demonstrates that early life experiences regulate methylation of the gene producing the glucocorticoid receptor, which results in lifelong changes to glucocorticoid secretion in response to challenges [cite]. A growing number of studies also demonstrate early life programming of DNA methylation patterns in wild populations resulting from dominance hierarchies [@laubach2019], brood size [@jimeno2019], temperature [cite], or landscape features [cite].